% Code

## Load packages

```{r message=FALSE, warning=FALSE}
library(readxl)
library(tidyverse)
library(ggthemes)
library(janitor)
library(patchwork)
```



## Customize ggplot themes

```{r}
my_theme <- theme_bw() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(color = "black", size = 0.5),
    axis.line.y = element_line(color = "white", size = 1),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank())

my_theme2 <- theme_bw() +
  theme(
    legend.position = "none",
    axis.line.x = element_line(color = "black", size = 0.5),
    axis.line.y = element_line(color = "white", size = 1),
    panel.border = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
  axis.text.y = element_text(face = "italic"))
  
my_theme3 <- theme_bw() +
  theme(
    legend.position = "none",
   axis.line.x = element_line(color = "black", size = 0.5),
    axis.line.y = element_line(color = "white", size = 1),
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
   legend.text = element_text(face = "italic"))

```


## Read data from XLSX

```{r}
fus <- read_excel("data/fusarium-parana.xlsx")
# using read_excel to deal better with encoding format
```

## Summary and visualization

### All isolates per host

```{r}
fus %>% 
  tabyl(Host)
```


### All strains per year

```{r}
fus %>% 
  tabyl(Host)
```


```{r}
Fig1A <- fus %>% 
  tabyl(Host) %>% 
    ggplot(aes(reorder(Host, n), n, label = n))+
     geom_col(fill = "#ffcc99", color = "black", size = 0.5)+
  theme(axis.text.x=element_text(face="italic"))+
      geom_text(position = position_dodge(width = 1),
            hjust = 0.5, vjust = -0.2) +
  labs(x = "", y = "Number of strains")+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 600)) +
  my_theme


```


### All strains per year

```{r}
fus %>% 
  tabyl(Year)
```

```{r}
Fig1B <- fus %>% 
  tabyl(Year) %>% 
    ggplot(aes(Year,n, label = n))+
    geom_col(fill = "#ffcc99", color = "black", size = 0.5)+
  theme_light()+
    geom_text(hjust = 0.5, vjust = -0.2) +
  labs(x = "", y = "Number of strains")+
    scale_y_continuous(expand = c(0, 0), limits = c(0, 350)) +
  my_theme

```


### All strains per complexes

```{r}
fus %>% 
  tabyl(Complexes)
```


```{r}
Fig1C <- fus %>% 
  tabyl(Complexes) %>% 
    ggplot(aes(reorder(Complexes, n),n, label = n))+
    geom_col(fill = "#ffcc99", color = "black", size = 0.5)+
   coord_flip()+
  theme(axis.text.x=element_text(face="italic"))+
      geom_text(position = position_dodge(width = 1),
            hjust = 0) +
  labs(x = "", y = "Number of strains")+
  scale_y_continuous( limits = c(0, 650)) +
  my_theme2

```

### Figure 1 

```{r}
fig1 <- ((Fig1A / Fig1B) | Fig1C) + plot_annotation(tag_levels = 'A')
ggsave("figs/fig1.png", width =6, height =4)
fig1

```


### All species by host 

##### Wheat

```{r}
fus %>% 
  filter(Host == "Wheat") %>% 
  tabyl(Complex2,  Year) 
```


```{r}
Fig2A <- fus %>% 
  filter(Host == "Wheat") %>% 
  tabyl(Complex2,  Year) %>% 
  gather(year, n, 2:6) %>% 
  ggplot(aes(year, n, fill = Complex2))+
    geom_col()+
scale_fill_brewer(palette = "Paired")+
 theme_light()+
  labs(x = "Year", y = "Number of strains", fill = "", title = "Wheat (n = 366)")+
   scale_y_continuous(expand = c(0, 0), limits = c(0, 180)) +
  my_theme3

```


#### Barley

```{r}
fus %>% 
  filter(Host == "Barley") %>% 
  tabyl(Complex2,  Year) 
```



```{r}
Fig2B <- fus %>% 
   filter(Host == "Barley") %>% 
  tabyl(Complex2,  Year) %>% 
  gather(year, n, 2:6) %>% 
  ggplot(aes(year, n, fill = Complex2))+
    geom_col()+
  scale_fill_brewer(palette = "Paired")+
   theme_light()+
  labs(x = "Year", y = "Number of strains", fill = "", title = "Barley (n = 421)")+
       scale_y_continuous(expand = c(0, 0), limits = c(0, 180)) +
  my_theme3

```

### Figure 2


```{r}
patch <- Fig2A + Fig2B & theme(legend.position = "right") 
p1 <- patch + plot_layout(guides = "collect")+ plot_annotation(tag_levels = 'A')
ggsave("figs/fig2.png", width =7, height =4)
p1
```


### FGSC by host within year

#### Wheat

```{r}
Fig3A <- fus %>% 
  filter(Complex2 == "FGSC") %>% 
  filter(Host == "Wheat") %>% 
  tabyl(Species,  Year) %>% 
  gather(year, n, 2:6) %>% 
  ggplot(aes(year, n, fill = Species, label = n))+
    geom_col()+
  scale_fill_brewer(palette = "Dark2")+
 theme_light()+
  labs(x = "", y = "Number of strains", fill = "FGSC", title = "Wheat (n = 266)")+
         scale_y_continuous(expand = c(0, 0), limits = c(0,130)) +
  my_theme3
```

#### Barley


```{r}
Fig3B <- fus %>% 
    filter(Complex2 == "FGSC") %>% 
   filter(Host == "Barley") %>% 
  tabyl(Species,  Year) %>% 
  gather(year, n, 2:6) %>% 
  ggplot(aes(year, n, fill = Species))+
    geom_col()+
  scale_fill_brewer(palette = "Dark2")+
   theme_light()+
  labs(x = "", y = "Number of strains", fill = "FGSC", title = "Barley (n = 305)")+
       scale_y_continuous(expand = c(0, 0)) +
  my_theme3

```

#### Figure 3


```{r}
patch1 <- Fig3A + Fig3B + theme(legend.position = "right") 
Fig3AB <- patch1 + plot_layout(guides = "collect")+ plot_annotation(tag_levels = 'A')  
ggsave("figs/fig3.png", width =8, height =4)
Fig3AB

```

### FTSC by host within year

#### Wheat
```{r}
Fig3C <- fus %>% 
    filter(Complex2 == "FTSC") %>%
   filter(Host == "Wheat") %>% 
  filter(Species != "F. acuminatum") %>% 
  tabyl(Species,  Year) %>% 
  gather(year, n, 2:6) %>% 
  ggplot(aes(year, n, fill = Species))+
    geom_col()+
    scale_fill_brewer(palette = "Set2")+
   theme_light()+
 
  labs(x = "Year", y = "Number of strains", fill = "FTSC", title = "Wheat (n = 25)")+
   scale_y_continuous(expand = c(0, 0), limits = c(0,12)) +
  my_theme3
```

#### Barley

```{r}
Fig3D <- fus %>% 
    filter(Complex2 == "FTSC") %>% 
   filter(Host == "Barley") %>% 
  tabyl(Species,  Year) %>% 
  gather(year, n, 2:6) %>% 
  ggplot(aes(year, n, fill = Species))+
    geom_col()+
  scale_fill_brewer(palette = "Set2")+
   theme_light()+
  labs(x = "Year", y = "Number of strains", fill = "FTSC", title = "Barley (n = 36)")+
     scale_y_continuous(expand = c(0, 0), limits = c(0,12)) +
  my_theme3
```

#### Figure 4

```{r}
patch2 <- Fig3C + Fig3D + theme(legend.position = "right") 
Fig3CD <- patch2 + plot_layout(guides = "collect")+ plot_annotation(tag_levels = 'A') 
ggsave("figs/fig4.png", width =8, height =4)
Fig3CD

```


## Independence tests

### Host vs. complexes

```{r}
host_complexes <- fus %>% 
  tabyl(Complexes, Host)

chisq.test(host_complexes)
fisher.test(host_complexes)

```

### year vs. complexes

```{r}
year_complex2 <- fus %>% 
  tabyl(Complex2, Year)

fisher.test(year_complex2, simulate.p.value = TRUE)

```

### FGSC vs. host

```{r}
host_fgsc <- fus %>% 
  filter(Complex2 == "FGSC") %>% 
  tabyl(Species, Host)
chisq.test(host_fgsc)

```

### FGSC vs. year
```{r}
year_fgsc <- fus %>% 
  filter(Complex2 == "FGSC") %>% 
  tabyl(Species, Year)
fisher.test(year_fgsc, simulate.p.value = T)

```



## Toxin data

```{r}
tox <- read_excel("data/toxin-ftsc.xlsx",1) %>% 
  gather(toxin, value, 3:7 )
tox

```
